---
title: "Trends Across Pokemon Generations"
subtitle: "INFO 526 - Summer 2024 - Final Project"
author: 
  - name: "Luis Hernandez"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "This project attempts to analyze the trends of certain characteristics of Pokemon across generations 1 through 7"
format:
   html:
    code-tools: true
    code-overflow: wrap
    embed-resources: true
editor: visual
execute:
  warning: false
  echo: false
---

## Introduction

The dataset used for this project is the following dataset: https://github.com/williamorim/pokemon/

This Pokemon dataset contains 22 columns of data and 949 rows containing data of Pokemon up to the 7th generation Pokemon games. The 7th generation of Pokemon includes Pokemon Sun & Moon, Pokemon Ultra Sun & Moon, and Pokemon Let's Go Pikachu & Let's Go Eevee. This dataset in particular only contains the statistics of Pokemon up to the games Pokemon Sun & Moon, so none of the new Pokemon introduced in Ultra Sun & Moon are present. The data set contains information regarding several different statistics of Pokemon. This includes: Name, Species ID, Height, Weight, Base EXP, Type Info, HP, Attack, Defense, Special Attack, Speed, Colors, Egg Groups, and Generation ID.

Using this dataset, we will attempt to answer the following questions.

## Question 1: Type Counts Across Generations

**Introduction:**

An important characteristic of Pokemon is their typing. Pokemon types are properties applied to Pokemon that determine a variety of characteristics for Pokemon, such as moves they can use most effectively or what moves they're weak against. The question I'd like to answer regarding Pokemon types is: How are types distributed for non Legendary/Mythical Pokemon and Legendary/Mythical Pokemon for Pokemon generations 1 - 7?

To answer this question I need two separate data frames with one having non Legendary/Mythical Pokemon and the other having only Legendary/Mythical Pokemon. For both data frames, I then need a count of the occurrance of every Pokemon type per generation. I'm interested in this question as I feel that it would be interesting to see if certain types have risen or fallen in popularity, or if the developers of Pokemon, Game Freak, have a bias towards certain types.

**Approach:**

In order to answer this question, I will be using a bar chart to get a summation of types of the whole dataset, and connected line plots for non-Legendary/Mythical Pokemon and Legendary/Mythical Pokemon types across generations. The bar chart will allow us to easily see and compare how many times a particular type is counted in the whole data set. This will show how common or uncommon certain types are across the whole dataset. The connected line plot will serve a similar purpose, but it will allow us to compare by generation by setting the x axis as the generation, and the y axis as the number of times a type was counted in that generation.

**Analysis:**

## 0 - Pre Process and Setup

```{r setup cache=TRUE}
#| label: Pre Process

if (!require("pacman")) 
  install.packages("pacman")

# use this line for installing/loading
pacman::p_load(tidyverse,
               ggthemes,
               patchwork,
               ggiraph,
               pokemon)

# set theme for ggplot2
ggplot2::theme_set(ggplot2::theme_grey(base_size = 12))

# set figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 7, # 7" width
  fig.asp = 0.618, # the golden ratio
  fig.retina = 3, # dpi multiplier for displaying HTML output on retina
  fig.align = "center", # center align figures
  dpi = 300 # higher dpi, sharper image
)

pokemon_data <- pokemon

pokemon_data <- pokemon_data |>
  rename(generation = generation_id)

#Trimming specific columns that will not be needed for the analysis
pokemon_data$id <- NULL
pokemon_data$species_id <- NULL
pokemon_data$base_experience <- NULL

pokemon_data$egg_group_1 <- NULL
pokemon_data$egg_group_2 <- NULL

pokemon_data$color_1 <- NULL
pokemon_data$color_2 <- NULL
pokemon_data$color_f <- NULL

pokemon_data$url_icon <- NULL
pokemon_data$url_image <- NULL

#Removing Pokemon without a generation ID
#Referenced https://tidyr.tidyverse.org/reference/drop_na.html

pokemon_data <- pokemon_data |>
  drop_na(generation)

#Referenced: https://infopoke.net/pokedex/legendary
#Data does not include indicator for legendary/mythical status. 
#This means I need to manually sort which Pokemon are legendary/mythical
pokemon_data <- pokemon_data |>
  mutate(
    legend_stat = case_when(
      row_number() %in% 
        c(144, 145, 146, 150, 151, 243, 
          244, 245, 249, 250, 251, 377, 
          378, 379, 380, 381, 382, 383, 
          384, 385, 386, 480, 481, 482, 
          483, 484, 485, 486, 487, 488, 
          489, 490, 491, 492, 493, 494, 
          638, 639, 640, 641, 642, 643, 
          644, 645, 646, 647, 648, 649, 
          716, 717, 718, 719, 720, 721, 
          785, 786, 787, 788, 791, 792, 
          800, 801, 802 ) ~ TRUE,
          TRUE  ~ FALSE
    ),
    
    total_stats = rowSums(across(c(hp, attack, defense, special_attack, special_defense, speed)))
    
  )

special_pokemon <- pokemon_data |>
  filter(legend_stat == TRUE)

regular_pokemon <- pokemon_data |>
  filter(legend_stat == FALSE)

type_colors <-  
  c(
    "normal" = "gray50",
    "fighting" = "darkorange",
    "flying" = "lightskyblue3",
    "poison" = "mediumpurple",
    "ground" = "sienna",
    "rock" = "tan",
    "bug" = "darkolivegreen3",
    "ghost" = "purple4",
    "steel" = "slategray",
    "fire" = "red3",
    "water" = "dodgerblue3",
    "grass" = "seagreen4",
    "electric" = "goldenrod",
    "psychic" = "deeppink3",
    "ice" = "cyan3",
    "dragon" = "steelblue3",
    "dark" = "grey30",
    "fairy" = "plum3"
  )


```

## 1 - Type Count Bar Chart

```{r}
#| label: Type Code

#Need to account for secondary typing of Pokemon, so we're pivoting both type columns into one
pokemon_data_longer <- pokemon_data |>
  pivot_longer(
    cols = starts_with("type_"),
    values_to = "types"
  ) |>
#Some Pokemon only have a single type, which causes NA values to show up when the data is pivoted
  drop_na(types)

pokemon_data_longer |> ggplot(aes(y =  fct_rev(fct_infreq(types)), fill = types)) +
  geom_bar() +
#Referenced: https://stackoverflow.com/questions/63653351/how-to-use-stat-count-to-label-a-bar-chart-with-counts-or-percentages-in-ggplo
  scale_fill_manual(values = type_colors) +
  geom_text(
    aes(label = after_stat(count)),
    stat = "count", hjust = 1.1, size = 3,
    color = "white"
  ) +
  labs(
    title = "Total Number of Pokemon Type Occurrences",
    subtitle = "Data up to Pokemon Sun & Moon (2016)",
    caption = "Data source: https://github.com/williamorim/pokemon/",
    x = "Type Count",
    y = "Type"
  ) +
  theme(
    legend.position = "none",
  )
```

## 2 - Type Count Per Generation

```{r}
#| label: Type Code 2

#Need to account for secondary typing of Pokemon, so we're pivoting both type columns into one
regular_pokemon_data_longer <- regular_pokemon |>
  pivot_longer(
    cols = starts_with("type_"),
    values_to = "types"
  ) |>
#Some Pokemon only have a single type, which causes NA values to show up when the data is pivoted
  drop_na(types)

#Getting numerical value for every type per generation
regular_type_count  <- regular_pokemon_data_longer |>
  count(generation, types)
#-------------------------------------------------------------------------------------------------------#
#Legendary/Mythical Pokemon

special_pokemon_data_longer <- special_pokemon |>
  pivot_longer(
    cols = starts_with("type_"),
    values_to = "types"
  ) |>
  drop_na(types)

special_type_count  <- special_pokemon_data_longer |>
  count(generation, types)


#-------------------------------------------------------------------------------------------------------#
#Referenced: https://r-graph-gallery.com/412-customize-css-in-interactive-ggiraph.html
#Referenced: https://ardata.fr/ggiraph-book/
#Used to make the line graph interactive

plot1 <- regular_type_count |> ggplot(
  aes(x = generation, 
      y = n,
      color = types,
      group = types,
      tooltip = types,
      data_id = types,
      )
  ) +
  geom_point_interactive(size = 1) +
  geom_line_interactive(linewidth = 0.5, alpha = 0.5) +
  scale_color_manual_interactive(
    values = type_colors) +
  scale_x_continuous(breaks = regular_type_count$generation) +
  labs(
    title = "Total Number of Type Occurrences by Generation\nfor non-Legendary/Mythical Pokemon",
    subtitle = "Data up to Pokemon Sun & Moon (2016)",
    caption = "Data source: https://github.com/williamorim/pokemon/",
    x = "Generation",
    y = "Type Count"
  ) +
  theme(
    legend.position = "right",
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10),
  ) + 
  guides(
     color = guide_legend(override.aes = list(size = 1), ncol = 2) 
  )

#-------------------------------------------------------------------------------------------------------#

plot2 <- special_type_count |> ggplot(
  aes(x = generation, 
      y = n,
      color = types,
      group = types,
      tooltip = types,
      data_id = types,
      )
  ) +
  geom_point_interactive(size = 1) +
  geom_line_interactive(linewidth = 0.5, alpha = 0.75) +
  scale_color_manual_interactive(
    values = type_colors) +
  scale_x_continuous(breaks = special_type_count$generation) +
  labs(
    title = "Total Number of Type Occurrences by Generation\nfor Legendary/Mythical Pokemon",
    subtitle = "Data up to Pokemon Sun & Moon (2016)",
    caption = "Data source: https://github.com/williamorim/pokemon/",
    x = "Generation",
    y = "Type Count"
  ) +
  theme(
    legend.position = "right",
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10),
  ) + 
  guides(
     color = guide_legend(override.aes = list(size = 1), ncol = 2) 
  )

#-------------------------------------------------------------------------------------------------------#

#This allows the graph to highlight the lines when hovered over along with hiding lines when clicking on the legend
interactive_type_plot1 <- girafe(ggobj = plot1)
interactive_type_plot2 <- girafe(ggobj = plot2)

interactive_type_plot1 <- girafe_options(
  interactive_type_plot1,
  opts_hover(css = "stroke-width: 3px; transition: all 0.3s ease;pointer-events: visibleStroke;"),
  opts_hover_key(css = "stroke-width: 3px; transition: all 0.3s ease;"),
  opts_hover_inv("opacity:0.15;filter:saturate(10%);"),
  opts_toolbar(saveaspng = FALSE)
)

interactive_type_plot2 <- girafe_options(
  interactive_type_plot2,
  opts_hover(css = "stroke-width: 3px; transition: all 0.3s ease;pointer-events: visibleStroke;"),
  opts_hover_key(css = "stroke-width: 3px; transition: all 0.3s ease;"),
  opts_hover_inv("opacity:0.15;filter:saturate(10%);"),
  opts_toolbar(saveaspng = FALSE)
)


interactive_type_plot1
interactive_type_plot2
```

**Analysis:**

##Question 2: Statistical and Physical Averages Across Generations
