---
title: "Proposal title"
subtitle: "Proposal"
author: "Luis Hernandez"
  - name: "Luis Hernandez"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
if (!require("pacman")) 
  install.packages("pacman")

# use this line for installing/loading
pacman::p_load(tidyverse,
               pokemon)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false

pokemon

```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

<https://github.com/williamorim/pokemon/>

The Pokemon dataset contains 22 columns of data and 949 rows containing data of Pokemon up to the 7th generation Pokemon games, which includes Pokemon Sun/Moon, Pokemon Ultra Sun/Moon, and Pokemon Let's Go Pikachua/Eevee. The initial com

The columns contain information regarding several different statistics of Pokemon. This includes: Name, Species ID, Height, Weight, Base EXP, Type Info, HP, Attack, Defense, Special Attack, Speed, Colors, Egg Groups, and Generation ID.

I chose this dataset because I'm interested in Pokemon, and as a series that has existed since the 90's there is plenty of data to sift through that may reveal trends across different generations.

## Questions

What is the type distributions between generations and does this show any trends that maybe indicate certain types becoming more or less popular?\
\
Have there been trends with the base stats of Pokemon between the newer and older generations such as rising base stats, normal base stats, or falling base stats?

## Analysis plan

- For the type data I would want to look at the number of Pokemon and types per generation and compare them across the generations in a bar. Pokemon with multliple types could either be counted in all their relevant types, or they could split off from singular type Pokemon into a different category consisting of Pokemon with multiple types.
- For the stats I need to look at the base stats and do data wrangling to figure out how to analyze stats across generations. My current idea is either get averages of stats within a single generation and then compare that average across generations.
