---
title: "Proposal title"
subtitle: "Proposal"
author:
  - name: Luis Hernandez
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
if (!require("pacman")) 
  install.packages("pacman")

# use this line for installing/loading
pacman::p_load(tidyverse,
               pokemon)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false

glimpse(pokemon)

```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

<https://github.com/williamorim/pokemon/>

The Pokemon dataset contains 22 columns of data and 949 rows containing data of Pokemon up to the 7th generation Pokemon games, which includes Pokemon Sun/Moon, Pokemon Ultra Sun/Moon, and Pokemon Let's Go Pikachu/Eevee.

The columns contain information regarding several different statistics of Pokemon. This includes: Name, Species ID, Height, Weight, Base EXP, Type Info, HP, Attack, Defense, Special Attack, Speed, Colors, Egg Groups, and Generation ID.

I chose this dataset because I'm interested in Pokemon, and as a series that has existed since the 90's there is plenty of data to sift through that may reveal trends across different generations.

## Questions

What is the type distributions between generations and does this show any trends that maybe indicate certain types becoming more or less popular?\
\
Have there been trends with the base stats of Pokemon between the newer and older generations such as rising base stats or falling base stats?

## Analysis plan

- For the type data I would want to look at the number of Pokemon and types per generation and compare them across the generations in a bar graph. Pokemon with multiple types will be counted in all their relevant types, though a stretch goal could be to split them off from singular type Pokemon into a different category consisting of Pokemon with multiple types. The y axis would be the number of pokemon, the x axis is the typing, and the bar fill is the generation. From here I would used Facet_Grid to display the different plots, and then compare the data per generation that way.
- For the stats I need to look at the base stats(HP, Attack, Defense, Special Attack, Speed) and do data wrangling to analyze stats across generations. My current idea is get averages of all the stats within a single generation and then compare that average across generations by utilizing line graphs for each stat where the y axis is the numerical value, and the x axis are the different generations, with each stat represented by a line.
- Some Pokemon in the dataset have no associated generation ID due to being variants of existing Pokemon. These variants will be cleaned from the dataset as these are not distinct Pokemon. Alongside these some columns will be dropped as they are not needed for this analysis (colors, egg groups, urls).

| Task Name | Status | Due | Priority | Summary |
|----|----|----|----|----|
| Proposal | Incomplete | June 17th | High | Need to finalize details of the project proposal |
| Final Deliverables | Incomplete | July 3rd | High | Final project items: Reproducibility, Presentation, Write-up |
| Stat Graph | Incomplete | June 20th | High | Need to finish all related work to the stat graph. Starting early on this graph as I predict it will take the most work |
| Type Graph | Incomplete | June 24th | High | All related work to Type Graph |
| Reproducibility | Incomplete | June 26th | Low |  |
| Presentation | Incomplete | June 28th | Medium |  |
| Write-Up | Incomplete | June 30th | Medium |  |

: Weekly Project Plan, Due dates are rough estimates of when task should be done.
